---
title: "NESARC Odds Ratio Analysis 222 Sample"
author: "Angel Garcia de la Garza"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output:
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js
---


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, 
                      results = 'asis',
                      cache = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      cache.lazy = FALSE)

library(pROC)
library(doParallel)
library(beepr)
#library(summarytools)
library(psych)
library(knitr)
library(forcats)
library(furrr)
library(randomForest)
library(gbm)
library(e1071)
library(caret)
library(glmnet)
library(tidyverse)



```

## R Markdown


```{r load_data, cache = TRUE}

load("/Users/angelgar/NESARC/Data/w1w2suicidedata.RData")
load("/Users/angelgar/NESARC/Data/w1w2suicidedata_WithCases_06_04_2019.RData")

```


```{r}

w1vars1[dim(w1vars1)[1] + 1, ] <- c("INCPER3", "Personal Income in last Year")
w1vars1[dim(w1vars1)[1] + 1, ] <- c("INCFAM3", "Family Income in last Year")

```


# Remove previous attempters

```{r process_data}

w1w2suicide <- w1w2suicide.v3

rm(w1w2suicide.v3)

w1w2suicide <- w1w2suicide %>%
                mutate(suicide_sli = suicide_new) %>%
                select(-W2S4AQ3A16,
                       -W2S14Q17A,
                       -W2S14Q17B,
                       -W2S14Q17C,
                       -suicide_life,
                       -suicide_age,
                       -age_diff,
                       -suicide_new)

```



# Step 1: Sort out numeric from non numeric

I created a decision tree that uses two flag the numeric variables. 

* All those variables that include number in either it's name or description
* All those variables that have more than 20 unique values

## Find those variables in the code book with "number"

```{r clean_values}

w1vars1 <- w1vars1 %>%
            mutate(num.description = grepl("num", w1vars1$description, ignore.case = TRUE),
                   num.description = grepl("scale", w1vars1$description, ignore.case = TRUE),
                   num.name = grepl("num", w1vars1$code, ignore.case = TRUE),
                   num.logical = num.description | num.name)
```

## Find variables with more than 20 unique values

```{r}

unique_values <- function(x) {
  length(unique(x))
}

w1w2_datatype <- w1w2suicide %>% 
                  map_dfc(~unique_values(.)) %>%
                  gather(key = "code", "unique") %>%
                  mutate(cont_twenty_logic = unique >19)

```

## Merge these two together to create one single rule

```{r}

w1w2_datatype <- w1w2_datatype %>%
                  left_join(w1vars1, by = "code") %>%
                  mutate(is_numeric = cont_twenty_logic | num.logical) %>%
                  select(code, description, is_numeric) %>%
                  unique() %>%
                  mutate(is_numeric = ifelse(is.na(is_numeric), FALSE, is_numeric)) %>%
                  mutate(is_numeric = ifelse(code == "S1Q1E", FALSE, is_numeric),
                         is_numeric = ifelse(code == "WEIGHT", FALSE, is_numeric),
                         is_numeric = ifelse(code == "STRATUM", FALSE, is_numeric),
                         is_numeric = ifelse(code == "PSU", FALSE, is_numeric),
                         is_numeric = ifelse(code == grepl("scale", 
                                                           description, 
                                                           ignore.case = T), FALSE, is_numeric)
                         ) %>%
                  select(code, is_numeric)

```


## Create Dummy Variables for Missing Continuous

```{r}

column.is.numeric <- w1w2_datatype$code[which(w1w2_datatype$is_numeric == TRUE)]


w1w2suicide_num <- w1w2suicide %>%
                    select(column.is.numeric)


numeric.tocat <- function(x) {
  
  dummy.unknown <- ifelse(x == 99, "unknown", NA)
  x.no.99 <- ifelse(x == 99, NA, x)
  
  quantiles <- ntile(x.no.99, 3) %>% as.character()
  quantiles <- ifelse(quantiles == "1", "low", 
          ifelse(quantiles == "2", "mid", "high"))
  
  quantiles <- ifelse(is.na(quantiles), dummy.unknown, quantiles)
  
  return(quantiles)

}


w1w2_numeric_quantiles <- w1w2suicide_num %>% 
                        map_dfc(~numeric.tocat(.))

names(w1w2_numeric_quantiles)[-1] <- paste(names(w1w2_numeric_quantiles)[-1], "quantiles", sep="_")
w1w2_numeric_quantiles$IDNUM <- w1w2suicide_num$IDNUM

```


## Retype categorical variables

```{r}

mutate.columns.tofactor <- w1w2_datatype$code[which(w1w2_datatype$is_numeric == FALSE)]

w1w2suicide <- w1w2suicide %>% 
                    select(IDNUM, mutate.columns.tofactor) %>%
                    left_join(w1w2_numeric_quantiles, by = "IDNUM") %>%
                    mutate_all(as.factor) %>%
                    mutate_all(fct_explicit_na)
                    

names(w1w2suicide)[-1] <- paste0(names(w1w2suicide)[-1], "_")

```


## Check for variables with unique values.

```{r}

remove.unique.cat <- w1w2suicide %>%
                      future_map_dfc(~length(unique(.))) %>%
                      t() %>%
                      as_tibble(rownames = "variable") %>%
                      rename(unique = V1) %>%
                      filter(unique == 1) %>%
                      .$variable

```

## Filter those out

```{r}

w1w2suicide <- w1w2suicide %>%
            dplyr::select(-IDNUM, -remove.unique.cat,
                   -STRATUM_, -PSU_,
                   -CDAY_quantiles_,
                   -CYEAR_,-CMON_,
                   -S1Q24FT_, -S1Q24IN_, -S1Q24LB_quantiles_,
                   -S1Q1E_)

```


## Section 6: Odds Ratio

```{r, cache = T}

plan(multiprocess)

logistic.model <- function(x) {
  
  model <- glm(suicide ~ x, family = "binomial")
  
  return(anova(model, test = "Chisq")$`Pr(>Chi)`[2])
  
}

set.seed(1)

suicide <- w1w2suicide$suicide_sli_

w1w2_OR <- w1w2suicide %>% 
                  select(-suicide_sli_, - WEIGHT_) %>%
                  future_map_dfc(~logistic.model(.)) %>%
                  t() %>%
                  as_tibble(rownames = "variable") %>%
                  mutate(z.stat = qnorm(V1, lower.tail = FALSE)) %>%
                  rename(p_value = V1) %>%
                  mutate(variable = str_replace(variable, "_",""))

```

## Section 6: Weighted Odds Ratio 


```{r, cache = T}

logistic.model.weighted <- function(x) {
  
  model <- glm(suicide ~ x, family = "binomial",
               weights = weight.nesarc)
  
  return(anova(model, test = "Chisq")$`Pr(>Chi)`[2])
  
}

weight.nesarc <- as.numeric(as.character(w1w2suicide$WEIGHT_))
weight.nesarc <- (weight.nesarc)/(sum(weight.nesarc))*34629

w1w2_OR_weighted <- w1w2suicide %>% 
                      select(-suicide_sli_, - WEIGHT_) %>%
                      future_map_dfc(~logistic.model.weighted(.)) %>%
                      t() %>%
                      as_tibble(rownames = "variable") %>%
                      mutate(z.stat = qnorm(V1, lower.tail = FALSE)) %>%
                      rename(p_value = V1) %>%
                      mutate(variable = str_replace(variable, "_",""))

```



## Final Table of Odds Ratio with weighted and unweighted variables.


```{r}

plan(sequential)

```


```{r}

w1w2_OR_all <- w1w2_OR_weighted %>%
                  rename(z.stat_weighted = z.stat,
                         p_value_weighted = p_value) %>%
                  left_join(w1w2_OR, by = "variable") %>%
                  mutate(rank_unweighted = rank(-z.stat),
                         rank_weighted = rank(-z.stat_weighted)) %>%
                  rename(z.stat_unweighted = z.stat,
                         p_value_unweighted = p_value)

```

## Final Table of Odds Ratio with weighted and unweighted variables.

```{r}

OR_Final <- w1w2_OR_all %>%
              mutate(variable = str_replace(variable, "quantiles_","")) %>%
              mutate(variable = str_replace(variable, "_","")) %>%
              rename(code = variable) %>%
              left_join(w1vars1, by = "code") %>%
              arrange(desc(z.stat_unweighted)) %>%
              select(code, description, z.stat_unweighted, 
                     z.stat_weighted, rank_unweighted, rank_weighted)
```

This analysis has the `r sum(as.numeric(as.character(suicide)))` cases.


```{r}

OR_Final %>%
  kable()

```


```{r}

write_csv(OR_Final, "/Users/angelgar/NESARC/Output/NESARC_OR_Output_222_06_04_2019.csv")

```

